Technique for \earlywarning\ filtering for compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-28 results.tex
Converted some figures to eps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1162 ccdd05d5-56c1-461a-bf98-85f6876523b4
tone down
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1170 ccdd05d5-56c1-461a-bf98-85f6876523b4

We examine the performance of the \lloid\ algorithm on a small region of
compact binary parameter space centered on typical \textsc{ns}--\textsc{ns}
masses.  We begin by constructing a template bank that spans component masses
between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
curve~\citep{ALIGONoise}.  This results in a grid of 98544 points, or
$2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
the parameter space by chirp mass.  We shall concentrate on a sub-bank with 657
chirp masses between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
Figure \ref{fig:tmpltbank} illustrates this procedure.
\begin{figure}[h]
	\plotone{figures/tmpltbank.pdf}
	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
case study, consisting of component masses $m_1$, $m_2$, between 1 and 3~$M_\odot$, and
chirp masses $\mathcal{M}$ between 1.1955 and 1.2045~$M_\odot$.}
\end{figure}
With this sub-bank we were able to construct an efficient time-slice decomposition
that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
in Table~\ref{tab:time_slices} and Figure~\ref{fig:envelope}.
\begin{table*}
\caption{\label{tab:time_slices} Filter design sub-bank of 657 templates.  From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.  We vary \SVD{} tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
\begin{center}
\begin{tabular}{rr@{,\,}lc*{6}{r}}
\tableline\tableline
\\ [-1.5ex]
\multicolumn{4}{c}{} &\multicolumn{6}{c}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\ [1ex]
\cline{5-10}
\\ [-1.5ex]
$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\ \tableline
4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
\tableline
\end{tabular}
\end{center}
\end{table*}
%
\begin{figure}[h]
	\plotone{figures/envelope.pdf}
	\caption{\label{fig:envelope}Amplitude envelope, time slices boundaries, and sample rates for sub-bank used for impulse response test.}
\end{figure}
%
We use this sub-bank and decomposition for the remainder of this section.

@@

\subsection{Lower bounds on computational cost compared to other methods}

@@

The 5~s latency is probably due to buffering and synchronization.  The latency could probably be reduced by carefully tuning buffer lengths at every stage in the pipeline.

>>>>

We examine the performance of the \lloid\ algorithm on a small region of
compact binary parameter space centered on typical \textsc{ns}--\textsc{ns}
masses.  We begin by constructing a template bank that spans component masses
between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
curve~\citep{ALIGONoise}.  This results in a grid of 98544 points, or
$2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
the parameter space by chirp mass.  We shall concentrate on a sub-bank with 657
chirp masses between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
Figure \ref{fig:tmpltbank} illustrates this procedure.
\begin{figure}[h]
	\plotone{figures/tmpltbank.pdf}
	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
case study, consisting of component masses $m_1$, $m_2$, between 1 and 3~$M_\odot$, and
chirp masses $\mathcal{M}$ between 1.1955 and 1.2045~$M_\odot$.}
\end{figure}
With this sub-bank we were able to construct an efficient time-slice decomposition
that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
in Table~\ref{tab:time_slices} and Figure~\ref{fig:envelope}.
\begin{table*}
\caption{\label{tab:time_slices} Filter design sub-bank of 657 templates.  From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.  We vary \SVD{} tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
\begin{center}
\begin{tabular}{rr@{,\,}lc*{6}{r}}
\tableline\tableline
\\ [-1.5ex]
\multicolumn{4}{c}{} &\multicolumn{6}{c}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\ [1ex]
\cline{5-10}
\\ [-1.5ex]
$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\ \tableline
4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
\tableline
\end{tabular}
\end{center}
\end{table*}
%
\begin{figure}[h]
	\plotone{figures/envelope}
	\caption{\label{fig:envelope}Amplitude envelope, time slices boundaries, and sample rates for sub-bank used for impulse response test.}
\end{figure}
%
We use this sub-bank and decomposition for the remainder of this section.

@@

\subsection{Lower bounds on computational cost and latency compared to other
methods}

@@

The 5~s latency is probably due to buffering and synchronization.  The latency
might be reduced by carefully tuning buffer lengths at every stage in
the pipeline.

Even without further refinements, our implementation of the \lloid\ algorithm
achieves latencies that would make other aspects of the analysis the limiting
steps assuming that the human vetting process is removed.
